### 2014-03-11 paper/paper.tex
Started the paper in LaTeX

\documentclass[12pt]{article}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage[margin=1in]{geometry}
\begin{document}

\begin{titlepage}
\title{Checkers with AI}
\date{March 14, 2014}
\author{Thomas White
    \and Nathan Rogers
    \and Daniel Shubin}
\maketitle
\end{titlepage}

\section{Introduction:}
    \subsection{Overview:}
        Our project was to create a checkers engine in Python,
        that is playable by two AI's and humans. The two AI's
        were written in C++ and Haskell.
    \subsection{Why we chose C++, Haskell, and Python:}
        We chose Python as the front end implementation language
        because Python allows you to easily interface with other
        languages. Python is also a scripting language.
        It is also a language that we knew and could efficiently
        implement the checkers engine.
        For our first AI, we chose C++ since our team has a lot of
        experience with the language, and is a good contrast to functional
        programming since it is purely imperative.
        For our second AI, we chose Haskell because we had experience from
        the class. And it is a contrast to the imperative languages since
        it is purely functional.

\end{document}